On the formation of H$\alpha$ line emission around classical T Tauri stars

### 2006-01-27 TTau_Halpha/kurosawa.tex
nearly final draft
git-svn-id: https://repository.astro.ex.ac.uk/torus/trunk@642 25464f21-6099-402c-b30a-640f1d40d38b

This type can be produced by the disc-wind-magnetosphere hybrid models
at moderate to high inclination angles with wide ranges of the wind
acceleration rate ($0.2 \lesssim \beta  \lesssim 2.0$) and the wind
temperature ($7000\,\mathrm{K} \lesssim T \lesssim 9000 \mathrm{K}$). If
the inclination is high ($i \sim 80^\circ$), the disc-wind
acceleration parameter should be relatively small ($\beta \lesssim
0.5$). If the inclination is moderate ($i \sim 60^\circ$), the wind
acceleration parameter should be relatively high ($\beta \gtrsim 2.0$).

@@

These

The physical conditions which result in this type
is very similar to those of Type~II-B according to the
disc-wind-magnetosphere hybrid models. The main difference is between  
the two is in their inclination angles. This type is more likely
seen at higher inclination angles ($i \sim 80^\circ$) compared to Type
II-B.  Commonly seen in the models with the disc wind only model.

@@

xxx describe this type xxxx

This type is the most commonly seen profiles in the CTTS H$\alpha$
samples in \citet{reipurth:1996}. Interestingly, the most common
profile type seen in the disc-wind-magnetosphere hybrid model are
Type~III-B profiles (c.f.~Figures~\ref{fig:atlas_hybrid} and
\ref{fig:atlas_hybrid}). In this type of profiles, the
main emission comes from the magnetosphere, and the profile shape is
altered from a slightly blue-asymmetric one to a slightly
red-asymmetric one via the wind absorption in the blue wing.  In the
models with the disc-wind, the acceleration parameter should be
relatively small ($\beta<0.5$). The profile type changes from Type
III-R to Type~II-R if the value of $\beta$ increases. The
inclination angles must be moderate to high ($ 60^\circ \lesssim i \lesssim
 80^\circ$).

@@

xxx describe this type xxxx

This is the least common type of the observed profile (2~per~cent)
according to \citet{reipurth:1996}. Incidentally, we also found it
difficult to reproduce this type of profile since this can be achieve
only at narrow range of parameter space (very high inclination angle
with a disc wind model, $T_{\mathrm{max}}\sim8300\,\mathrm{K}$, and
$\dot{M}_{\mathrm{acc}}\sim10^{-8}\,\mathrm{M_{\sun}yr^{-1}}$).
Since the inclination angles must be high, this
indicates that the accretion disc plays an important role in producing
this type of the profile, but not the wind.  The rareness of this type in
the observation can be understood by the required high inclination
angle (almost edge-on) in our models.

@@

xxx describe this type xxxx

This type of the profiles is very similar to Type III-B, but the wind
in this type of profile may be accelerating faster (a smaller $\beta$
value) than that in Type~III-B.  In the models with the disc-wind,
this type is seen at moderate to high inclinations ($60^\circ \lesssim
 i \lesssim  80^\circ$) with relatively small value of the acceleration
parameters ($\beta \lesssim 0.5$).

@@

xxx describe this type xxxx

This is the second least common type of the observed profiles (5~per~cent)
which shows the inverse P-Cyg profile shape.  This can be understood
since the magnetospheric accretion model of \citet{hartmann:1994}
predicts that the redshifted absorption can occur at a limited range
of inclination angle ($i\simeq 50^\circ$, but depends on the geometry of
the magnetosphere) with which the hot continuum flux from the
footprint of the magnetosphere is seen through relatively fast moving
gas in the accretion stream.  The wind component must be very
weak. The mass-accretion rate should be much lower
($\dot{M}_{\mathrm{acc}} \sim 10^{-9}\,\mathrm{M_{\sun}\,yr^{-1}}$)
than that of the models which produce Type~I profiles since the flux
in the broadening wings can not be too high to produce the inverse
P-Cyg profile.

@@

\subsection{Inclination dependency of H$\alpha$ profiles}

@@

  \caption{Dependency of the H$\alpha$ model profiles on the inclination
    angle. The profiles computed at $i=10^{\circ}$(solid), $55^{\circ}$
    (dot) and $80^{\circ}$ (dash) using (1)~the
    disc-wind-magnetosphere hybrid model (left;
    c.f. section~\ref{sub:result-discwind-knigge}) and (2)~the disc, stellar wind,
    magnetosphere hybrid model (right panel;
    c.f. Figure~\ref{fig:config_stellarwind}) are shown for a
    comparison.  The parameters used for the magnetosphere are same for
    the both models, i.e. $T_{\mathrm{max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}~\mathrm{M_{\sun}~yr^{-1}}$.  The
    isothermal wind temperature and the wind acceleration parameter used
    in (1) are $T_{\mathrm{wind}}=7000~\mathrm{K}$ and $\beta=1.0$,
    and those in (2)
    are $T_{\mathrm{wind}}=9000~\mathrm{K}$ and $\beta=0.5$ respectively.
    All the other parameters are same as in
    sections~\ref{sub:result-discwind-hybrid} and
    \ref{sub:result-discwind-knigge}. While the line equivalent width
    decreases as the inclination angle increases for the disc-wind
    model, it increases as the inclination angle increases for the
    stellar wind model. Only the latter agrees with the tendency seen in the
    observational study of \citet{appenzeller:2005} who showed that the
    H$\alpha$ line equivalent width of CTTS increases as the inclination
    angle increases. }

@@

In the previous sections, we have considered the H$\alpha$ line
formations in the disc wind which mimic the density distribution of the
magneto-centrifugal launched jet numerical models
(e.g.~\citealt{krasnopolsky:2003}). The density in this model is more
concentrated toward the equatorial plane in near field. 
In this section, we investigate the consistency of the line formation
in the disc-wind model with the recent the recent observation of
\citet{appenzeller:2005} by concentrating on the inclination
dependency of the EW of H$\alpha$.  Their study showed that 
that the line equivalent width of CTTS increases
as the inclination angle increases.

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the profiles computed with the disc-wind, magnetosphere hybrid model
(section~\ref{sub:result-discwind-knigge}) 
at inclination angles $i=10^{\circ}$, $55^{\circ}$ and
$80^{\circ}$. The magnetospheric accretion component uses the  
parameters used as the reference model (Table~\ref{tab:std_parameters}).  
The figure shows that the absorption
feature becomes stronger as the inclination increases, mainly because
of the geometrical configuration.  The optical depth to becomes larger
as the inclination angle increases since there is more material
towards the equatorial plane.  This tendency
of the inclination dependency holds for the models with the wide
ranges of $\beta$ and $T_{\mathrm{wind}}$ values.  
As a result, the EW becomes smaller as the inclination
angle increases: 41\AA, 27\AA, and 19\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively. 
Clearly, this is not consist with the trend seen in
the observation of \citet{appenzeller:2005}. Recall that a similar
inclination dependency was found for the models with only
magnetosphere (Figure~\ref{fig:acc_inclination_effect}).

We have also considered an alternative wind model which would produce
the inclination dependency of the H$\alpha$ EW  seen in the
observation. In this model, the disc-wind is replaced by the stellar
wind moving only in radial direction with the classical beta-velocity
law \citep[c.f.][]{castor:1979}
i.e.
$v_{r}\left(\ r  \right) = v_{\infty} \left( 1 - R_{*}/r
  \right)^{\beta}$  where  $v_{\infty}$ and  $R_{*}$ are the 
  wind terminal velocity and the stellar radius respectively. $\beta$
  is the wind acceleration parameter similar to the one seen in the
  disc-wind model (c.f.~equation~\ref{eq:discwind-poloidal-velocity}).
The wind emerges from the stellar surface, but restricted to be within
$\sim30^{\circ}$ from the symmetry axis hence forming the cone-shaped 
regions both above and below the poles. The density is computed from
assuming the constant mass-loss rate per unit area over the stellar
surface, the mass-flux conservation, and using the beta velocity low
above.  The schematic of the model configuration is shown in
Figure~\ref{fig:config_stellarwind}. Again, we assume the wind
temperature ($T_{\mathrm{wind}}$) is isothermal for simplicity.  
This model configuration is motivated by the recent study by
\citet{matt:2005} who demonstrated the possibility that the stellar
wind along the open magnetic field originated from the star can cause
significant spin-down torque on the star, provided that mass-loss rate
is high enough.

The right panel in Figure~\ref{fig:effect_inclination_compare} shows
the H$\alpha$ profiles computed using the alternative hybrid model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. The profiles show
rather complex absorption features because of the interaction of
photons created in the wind with the magnetosphere. 
While the blue-shifted absorption features are formed in the wind, the
absorption features near the line centre and in the red wing (only for
$i=10^\circ$) are caused by the wind emission by the magnetosphere
which has a rather complex flow geometry. 
Because of the geometry of the wind, the P-Cyg absorption feature
weakens as the inclination increases. The optical depth of the wind is
much higher in 
the polar direction compared to that at higher inclination. For the
same reason, the EW of the line increases as the inclination increase
as clearly seen in the figure. Specifically their values are 
49\AA, 60\AA, and 84\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively.  This tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ as long as
the emission from the wind is comparable to that of the magnetosphere.

The distinct difference in the density dependency on the polar angles
in two different wind models should be apparent in the inclination
angle dependency on the observed line profiles. For example, one
possible way to distinguish the two different wind models is to 
examine the profile shape of a CTTS which is known to be viewed at
pole-on. If the profile contains a prominent blue-shifted absorption
component (e.g. Type IV-B), the system is more likely to be associated
with the `bipolar' stellar wind type of outflow (but not with the disc-wind type
of outflow).  In reality, two types of the winds (bipolar stellar wind and
disc wind) may co-exits similar to that seen in
e.g. \citet*{drew:1998} and \citet{matt:2005}. In these models, the
fast stellar wind is present in the polar directions, and at the same
time the slower disc wind is present near the equatorial plane.

The extensive parameter studies of the simple stellar wind model
presented in this section have not been performed. The models
presented here are for a demonstrative purpose, and the further
investigation is necessary before the models are applied to a real
system, perhaps, by combining our radiative transfer model with the
results from MHD 'stellar' wind calculations.

In this analysis, we found only the models with the bipolar stellar
wind agree with the trend seen in their observation. Although the
uncertainty in the inclination angles of the observed object might be
large, a larger sample in addition to that of \citet{appenzeller:2005}
for further investigation of the inclination dependency on the line
profile will be extremely useful for determining the wind
configuration of CTTS.

@@

  \caption{The alternative hybrid model. The model replaces the
    disc-wind in the previous hybrid model with the stellar wind. The
    wind launching from the stellar surface is limited within the
    cones with the opening angle $\sim30^\circ$ to avoid the wind from
    overlapping with the magnetosphere.  The system consists of four
    components: (1)~the continuum source located at the origin $\left( o
    \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
    $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
    (3)~the stellar wind outflow, and (4)~the accretion disc. The density
    distribution is symmetric around the $z$-axis. }

@@

We have presented the disc-wind-magnetosphere hybrid radiative
transfer models for classical T~Tauri stars, and detailed studies of
the H$\alpha$ formation from their complex to circumstellar
environment.  We found that hybrid model can reproduce the wide
variety of profile seen in observations
(Figures~\ref{fig:atlas_acc_models}, \ref{fig:atlas_discwind_only},
\ref{fig:atlas_hybrid} and \ref{fig:atlas_reipurth}). 
Especially, the blue-shifted
wind absorption feature very commonly seen in the H$\alpha$ profile of
CTTS were well understood by this model.  This was not possible in the
magnetospherical accretion only model presented in earlier works
(e.g.~\citealt{muzerolle:2001}).

The profiles computed with the hybrid model showed that the wind
emission contribution becomes significantly larger than that from the
magnetosphere (Figure~\ref{fig:atlas_hybrid} and
Table~\ref{tab:ew_ratio}).  Henceforth, a large H$\alpha$ equivalent
width does not necessary indicate a large mass-accretion rate, but it
could be caused by the mass-loss by the wind outflow.

@@

The value of $\mu$, 0.1,  suggested by both MHD calculations and observations
(c.f.~\citealt{koenigl:2000}) is consistent with the 
disc-wind-magnetosphere models presented here. We found that the
hybrid models with $0.05<\mu<0.2$ are also consistent with the type
of the profiles seen in the observations (e.g~\citealt{reipurth:1996}).

Unfortunately, the dependency of the line equivalent width on
the inclination angle predicted by  
the disc-wind-magnetosphere hybrid model does not agree with the trend
seen the observation of \citet{appenzeller:2005}; however, their
results are based on a small number of objects (12). We have also
considered an alternative hybrid model in which the disc-wind is
replace by the narrow cone-shaped stellar wind, and found that the
alternative wind hybrid model can reasonably reproduces the 
line equivalent width dependency on the inclination seen in the
observation (Figure~\ref{fig:effect_inclination_compare}).

Similar model configurations used here may be also applicable to Herbig
Ae/Be and brown dwarfs since their H$\alpha$ observations also exhibit
evidence of the outflow and inflow (e.g.~\citet{finkenzeller:1984};
\citealt{muzerolle:2005}), but no model has been developed to explain
the phenomenon. Further investigation on the possibility of applying
our model to these objects are needed.

Future work should include the followings: (1)~Improving the wind
models. The parametrisation of the wind density structure should be
changed to follow more closely that of MHD simulations. The wind
temperature structure should be calculated self-consistently
(c.f.~\citealt{hartmann:1994}; \citealt{martin:1996}). (2)~Computing
the line profiles using the density structure from MHD directly. Check
the consistency, and give feedback to the MHD models which are needed
to be tested against observational data. (3)~Spectropolarimetric study
to explore the geometry and the rotation of the disc
(e.g.~\citealt{vink:2005}). (4)~Modelling of the extended H$\alpha$
and spectro-astrometric observations (c.f.~\citealt{takami:2003};
\citealt{appenzeller:2005}). (5)~Line variability study using a 3D
radiative transfer model (e.g.~\citealt{symington:2005};
\citealt{kurosawa:2005a}) to probe the geometrical structure.

@@

Authors thank Jorick Vink for providing us the H$\alpha$ data of
T~Tau. This work is supported by PPARC standard grand PPA/G/S/2001/00081.

@@

Additional H$\alpha$ model line profiles 
computed using disc-wind-magnetosphere hybrid model
(section~\ref{sub:result-discwind-hybrid}) are presented here.  All
the models use the reference parameters for the 
magnetospherical geometry and the photosphere as given in
Table~\ref{tab:std_paramters} in the main text. The ratio ($\mu$) of the
mass-loss rate to mass-accretion rate is fixed to 0.1 for the all 
models. Three sets of model profiles, each with a different
mass-accretion rate, are presented here: 
(1) $\dot{M}_{\mathrm{acc}}=10^{-7}\,M_{\sun}\,\mathrm{yr^{-1}}$ (Figure~\ref{fig:atlas01}), 
(2) $\dot{M}_{\mathrm{acc}}=10^{-8}\,M_{\sun}\,\mathrm{yr^{-1}}$ (Figure~\ref{fig:atlas01}), and
(3) $\dot{M}_{\mathrm{acc}}=10^{-7}\,M_{\sun}\,\mathrm{yr^{-1}}$
(Figure~\ref{fig:atlas01}).   The temperature of the magnetosphere
used for (1), (2) and (3) are $7500\Kelvin$, $8500\Kelvin$ and
$9500\Kelvin$ respectively.  The corresponding wind mass-loss rates
are used for $\mu=0.1$. In each set the profiles are presented for the
inclination angle of $10^\circ$, $55^\circ$, and $80^\circ$. 
The disc-wind geometry used to compute these profiles are same as in
section~\ref{sub:result-discwind-hybrid}.

@@

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-6}\,M_{\sun}\,\mathrm{yr{-1}}$. }

@@

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-8}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$. }

@@

  \caption{The H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-9}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-10}\,M_{\sun}\,\mathrm{yr{-1}}$. }

>>>>

We find that Type~II-B profiles only occur at medium to high inclinations, and for systems
with mass accretion rates of $\approx 10^{-7}\,
\dot{M_\odot}\,\mathrm{yr^{-1}}$. If the system is viewed at high inclination, a Type~II-B profile implies
a fast wind acceleration, and conversely a moderate inclination angle suggests
a more modest wind acceleration.

@@

These profiles are characterised by a secondary red peak that is in excess of half the strength
of the primary peak. They are approximately as common as the Type~II-B lines.

For the  $\dot{M_{\rm acc}} = 10^{-7}\,
\dot{M_\odot}\,\mathrm{yr^{-1}}$ a switch from Type~II-B profiles to Type~II-R occurs
when the inclination angle changes from $55^\circ$ to $80^\circ$. This profile type is 
also seen at lower mass-accretion rates, but typically only in the models with the
slowest accelerating winds. It is perhaps unsurprising that the Type~II-B and Type~II-R profiles
are equally common, since (in general) the split between the profile types
results from a geometrical (viewing angle) effect rather than a marked difference in physical
parameters (mass accretion rate, wind acceleration etc).

@@

Profiles with a blue secondary peak that is less than half the strength of the 
primary peak are classified as Type~III-B. This type of profile is the most common
in the Reipurth sample, comprising 14 out of 43 (33 per cent) CTTS.

We find Type~III-B profiles only in our  $\dot{M_{\rm acc}} = 10^{-7}\,
\dot{M_\odot}\,\mathrm{yr^{-1}}$ models. A fast wind acceleration and a moderate 
inclination are necessary, or a slower accelerating wind viewed pole-on. 
We do not see this profile type in our high inclination models.

@@

This profile type has a red secondary peak that is less than half the strength
of the primary peak, and only one object with this type of profile was observed in the 
Reipurth sample (SZ Cha).

We are able to reproduce this profile type, but only using a very narrow range of
parameter space. In fact there are no Type~III-R profiles in our model grid, but
by 'tweaking' the magnetospheric temperature, and using an inclination ($85^\circ$) sufficient
that the disc starts to obscure the central H$\alpha$ emission. For completeness, the final
model parameters we used to achieve the profiles are:  (very high inclination, $T_{\mathrm{max}}\sim8300\,\mathrm{K}$, and $\dot{M}_{\mathrm{acc}}\sim10^{-8}\,\mathrm{M_{\sun}yr^{-1}}$).

It is interesting that the most infrequently observed profile type was in fact the most difficult
for us to reproduce. It appears that this profile morphology requires some obscuration effects
by the dust disc, and therefore a high inclination; this naturally explains the rarity of the profile
in the observations.

@@

This  profile classification is the classical P-Cygni, in which the blue-shifted absorption component
has a sufficient velocity to be present beyond the emission line wing. It occurs infrequently
in observations (7 per cent of the Reipurth sample).

No Type~IV-B profiles were produced in our model grid, since the blue wing of the magnetosphere
has generally such a large extent that it always exceeds the projected wind terminal velocity, and
thus the bluemost extent of the absorption component. Only by running a model with 
a larger terminal velocity, were we able to achieve a classical P-Cygni shape. We suspect
that the disc-wind geometry is not the best kinematic model for producing this morphology. A more
straightfoward explanation lies with a bipolar flow, viewed along its axis, or even a simple spherical wind.

@@

This is the inverse P-Cygni profile type, and only occurs in about 5~per~cent of 
H$\alpha$ profiles.

Type~IV-R profiles occur very infrequently in our hybrid model grid, with only
the lowest mass-accretion rate models viewed at high inclination showing the 
characteristic shape (although using the EW criterion these models would not
be classified as CTTS). The inverse P-Cygni profile requires that the observer
view the hot accretion footprints through the fast moving magnetospheric 
accretion flow, and this means that the inclination angle is restricted to 
small range.

We can reproduce Type-IV-R profiles more readily if we adopt a cooler magnetospheric
temperature than indicated by the accretion-rate/temperature relationship presented
by Muzerolle (xxxx).

@@

\subsection{The inclination dependence of H$\alpha$ profiles}

@@

  \caption{The dependence of the H$\alpha$ model profiles on the inclination
    angle. The profiles computed at $i=10^{\circ}$(solid line), $55^{\circ}$
    (dotted line) and $80^{\circ}$ (dashed line ) using (1)~the
    disc-wind-magnetosphere hybrid model (left;
    c.f. section~\ref{sub:result-discwind-knigge}) and (2)~the disc, stellar wind,
    magnetosphere hybrid model (right panel;
    c.f. Figure~\ref{fig:config_stellarwind}).  The parameters used for the magnetosphere are same for
    the both models, i.e. $T_{\mathrm{max}}=7500~\mathrm{K}$ and
    $\dot{M}_{\mathrm{acc}}=10^{-7}~\mathrm{M_{\sun}~yr^{-1}}$.  The
    isothermal wind temperature and the wind acceleration parameter used
    in (1) are $T_{\mathrm{wind}}=7000~\mathrm{K}$ and $\beta=1.0$,
    and those in (2)
    are $T_{\mathrm{wind}}=9000~\mathrm{K}$ and $\beta=0.5$ respectively.
    All the other parameters are same as in
    sections~\ref{sub:result-discwind-hybrid} and
    \ref{sub:result-discwind-knigge}.  }

@@

Recently, \citet{appenzeller:2005} have shown that the EW of H$\alpha$ is inclination
dependent, with the EW increasing as the inclination angle increases. Is our hybrid 
model consistent with this observation?

The left-hand panel in Figure~\ref{fig:effect_inclination_compare} shows
the profiles computed with the disc-wind, magnetosphere hybrid model
(section~\ref{sub:result-discwind-knigge}) 
at inclination angles $i=10^{\circ}$, $55^{\circ}$ and
$80^{\circ}$. The magnetospheric accretion component uses the  
parameters used as the reference model (Table~\ref{tab:std_parameters}).  
The figure shows that the absorption
feature becomes stronger as the inclination increases, mainly because
of the geometrical configuration.  The optical depth to the observer becomes larger
as the inclination angle increases, since the density of the disc wind increases towards
the equatorial plane--as a result, the EW becomes smaller as the inclination
angle increases: 41\AA, 27\AA, and 19\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively.  This is a robust result, and is insensitive to the adopted wind 
parameters. Recall that this is the same EW/inclination dependency as the magnetosphere
only models (Figure~\ref{fig:acc_inclination_effect}).

Clearly the inclination dependency of the line strength is in contradiction with
the  \citet{appenzeller:2005} result, and thus merits further investigation. 
In the previous sections, we have considered H$\alpha$ line
formation in a disc wind, which mimics the density distribution of the
magneto-centrifugal launched jet  models
(e.g.~\citealt{krasnopolsky:2003}). The density in this model is more
concentrated toward the equatorial plane in the region where the line emissivity is highest, although further out the material becomes collimated into a bipolar jet this would not 
contribute significantly to the line emission in our models.

As an alternative, we consider a wind model in which the outflow density
increases towards the pole, rather than the  equator. In this model, 
the disc-wind is replaced by a stellar wind, propagating only in the radial
direction, accelerating under the  classical beta-velocity law \citep[c.f.][]{castor:1979}
i.e.
$v_{r}\left(\ r  \right) = v_{\infty} \left( 1 - R_{*}/r
  \right)^{\beta}$  where  $v_{\infty}$ and  $R_{*}$ are the 
  wind terminal velocity and the stellar radius respectively. $\beta$
  is the wind acceleration parameter similar to the one seen in the
  disc-wind model (c.f.~equation~\ref{eq:discwind-poloidal-velocity}).
The wind emerges from the stellar surface, but restricted to be within
$\sim30^{\circ}$ from the symmetry axis hence forming the cone-shaped 
regions both above and below the poles. The density is computed from
assuming the constant mass-loss rate per unit area over the stellar
surface, the mass-flux conservation, and using the beta velocity low
above.  The schematic of the model configuration is shown in
Figure~\ref{fig:config_stellarwind}. Again, we assume the wind
temperature ($T_{\mathrm{wind}}$) is isothermal for simplicity.  
This model configuration is motivated by the recent study by
\citet{matt:2005} who demonstrated that the stellar
wind along  open magnetic field lines originating from the star may cause
significant spin-down torque.

The right-hand panel in Figure~\ref{fig:effect_inclination_compare} shows
H$\alpha$ profiles computed using the alternative model at
the same inclination angles used for the disc-wind-magnetosphere
hybrid model. The wind acceleration parameter ($\beta$) and the wind
temperature ($T_{\mathrm{wind}}$)used here are $0.5$ and
$9000\,\Kelvin$ respectively. All the other parameters are same as in
the disc-wind magnetosphere model.  With this set of $\beta$ and
$T_{\mathrm{wind}}$, the wind emission is comparable or a slightly
greater than the emission from the magnetosphere. The profiles show
rather complex absorption features because of the interaction of
photons created in the wind with the magnetosphere.

xxx check this out and rephrase xxx

The blue-shifted absorption features are formed in the wind, the
absorption features near the line centre and in the red wing (only for
$i=10^\circ$) are caused by the wind absorbing emission from the  magnetosphere.

Because of the geometry of the wind, the P-Cygni absorption feature
weakens as the inclination increases: the optical depth of the wind is
much higher in 
the polar direction. For the
same reason, the EW of the line increases as the inclination increases
as clearly seen in the figure. Specifically the values are 
49\AA, 60\AA, and 84\AA\, for $i=10^\circ$, $55^\circ$, and
$80^\circ$ respectively.  This tendency
of the inclination dependency holds for the models with this wind
models for wide ranges of $\beta$ and $T_{\mathrm{wind}}$ providing that
the emission from the wind is comparable to that of the magnetosphere.

The \citet{appenzeller:2005} study used just xxx objects, and a significantly
greater sample should be investigated before firm  conclusions are
drawn. However our models indicate a significant difference in the 
inclination dependence of the line morphology which may be simply
tested against observation.  For example, one
possible way to distinguish the two different wind models is to 
examine the profile shape of a CTTS which is known to be viewed close
to pole-on (e.g. a CTTS with a very low $v \sin i$). 
If the profile contains a prominent blue-shifted absorption
component (e.g. Type IV-B), the system is more likely to be associated
with the bipolar stellar wind type of outflow, and not with the disc-wind.
In reality, however, the two types of winds may be present in the same 
object, in a similar fashion to that seen in the models of 
\citet*{drew:1998} and \citet{matt:2005}. In these models, the
fast stellar wind is present in the polar directions, while at the same
time a slower, denser disc-wind is present near the equatorial plane.

@@

  \caption{The alternative outflow model. This model replaces the
    disc-wind with a stellar wind. The
    wind is launched from the stellar surface, and is limited to 
    cones with opening angles of $\sim30^\circ$.  The system consists of four
    components: (1)~the continuum source located at the origin $\left( o
    \right)$ of the cartesian coordinates $\left(x,y,z\right)$ -- the
    $y$-axis is into the paper, (2)~the magnetospheric accretion flow,
    (3)~the stellar wind outflow, and (4)~the accretion disc. The density
    distribution is symmetric around the $z$-axis. }

@@

We have presented disc-wind-magnetosphere hybrid radiative
transfer models for classical T~Tauri stars, and detailed studies of
the H$\alpha$ formation from their complex to circumstellar
environment.  We found that the hybrid model can reproduce the wide
variety of profile seen in observations
(Figures~\ref{fig:atlas_acc_models}, \ref{fig:atlas_discwind_only},
\ref{fig:atlas_hybrid} and \ref{fig:atlas_reipurth}).

@@

One potential free parameter of our models is the ratio of mass-loss rate to mass-accretion rate
($\mu$). Most MHD simulations predict a value of  $\mu = 0.1$ (c.f.~\citealt{koenigl:2000}), 
and we have adopted this figure for most of our models. The success of our
profile in matching observations is therefore encouraging,  and provides support
for this canonical value of $\mu$. We have computed sample models with other
values of $\mu$, (cite figure xxxx) and find that models with $0.05 <\mu<0.2$ are still
be consistent with observations. Wind-only or magnetosphere-only models may be
representative of individual line profiles, but neither can explain the complete
Reipurth classification scheme--even magnetosphere dominated profiles (Type~I)
can be readily explained by the hybrid model with $\mu=0.1$.

Unfortunately, the dependency of the line equivalent width on
the inclination angle predicted by  
the disc-wind-magnetosphere hybrid model does not agree with the trend
seen the observations of \citet{appenzeller:2005}. We may appeal
to the small sample size of their study (12 objects) to question
the validity of the EW effect, but we have have also
considered an alternative  model in which the disc-wind is
replaced by a biconical stellar wind, and have found that this
alternative wind model can reproduce the 
line equivalent width dependency on the inclination 
(Figure~\ref{fig:effect_inclination_compare}).  However, MHD simulations
predict that bipolar or jet-like flows in CTTS are always accompanied by
a disc-wind originating near the star: the real situation therefore is likely
to be a combination of the two flows.

Although we have concentrated on CTTS here, the hybrid model is equally
applicable to pre-main-sequence stars across the mass spectrum. Herbig
Ae/Be star profiles are equally diverse, and whilst many of the profiles
are indicative of a wind (\citet{finkenzeller:1984}) some lines may be explained
by magnetospheric accretion alone (e.g. \citealt{muzerolle:2005}). One possible
advantage in applying the hybrid model to the Herbig Ae/Be systems is the growing
observational data probing the line formation region and inner disc using techniques
such as spectro-astrometry (Baines xxxx), spectropolarimetry (Vink xxxx) and
interferometry (Monnier etc). These high-resolution techniques may aid model
fitting of individual objects by constraining some of the model parameters, in
particular the inclination.

Velocity broadened, asymmetric H$\alpha$ profiles in BDs are used as an
indicator of ongoing magnetospheric accretion, and the same models that
are used for CTTS are applied to these systems in order to determine mass-accretion
rates (muzerolle, lawson, jayawardhana xxxx). The line profiles however display
the same wide variety of morphologies as the CTTS and Herbig Ae/Be stars (although
we note as yet no BD profile has been observed with red-shifted absorption).
It is natural to assume that the same physics applies in these low mass objects, 
and that a hybrid model may be more successful in fitting these profiles, and will
give greater insight into the circumstellar kinematics.

There are several avenues of future work to consider. Here we have adopted
a simple analytical form for the accretion and outflow velocity and density structure.
A more self-consistent approach would be to use the results of MHD simulations
as an input to the radiative-transfer modelling. The thermodynamics of the 
circumstellar gas needs to be examined (c.f.~\citealt{hartmann:1994}; \citealt{martin:1996})
and also computed self-consistently with the radiation field used in the radiative-transfer,
a process that would decrease the free parameters of the model.
We intend to calculate other observables from the model that may be compared with observation 
and should provide further constraints. The line polarization signature, which
arises from scattering of line and continuum photons in the circumstellar dust, encodes both dynamical and geometric information, such as the size of the inner disk radius  (Vink obs paper, Vink theory paper xxxx). Spectro-astrometry may provide information on the system's inclination, and the geometry
of the outflow on large scales (Takami paper; Baines paper xxxx). Interferometry also provides information on inner disc radii and inclinations, although currently only for the Herbig Ae/Be stars.

We have concentrated on the formation of the H$\alpha$ line, since it is the best studied CTTS
emission line. However it is clear that simultaneous fitting of several emission line profile should
provide much stronger constrains. In particular it appears that the near-IR hydrogen lines (Pa$\beta$ and Br$\gamma$) are less-effected by outflows than H$\alpha$. One could therefore use these
lines to constrain the mass-accretion rate, and H$\alpha$ the mass-loss rate. A potential pitfall 
is the significant temporal variability that all CTTS show in their emission lines; the observational
data on a particular object must be co-temporaneous.

@@

This work is supported by PPARC standard grand PPA/G/S/2001/00081
and PPARC rolling grant xxxxx.

@@

Here we present our H$\alpha$ model line profile grid 
computed using disc-wind-magnetosphere hybrid model
(section~\ref{sub:result-discwind-hybrid}).  All
the models use the reference parameters for the 
magnetospheric geometry and the photosphere as given in
Table~\ref{tab:std_paramters} in the main text. The ratio ($\mu$) of the
mass-loss rate to mass-accretion rate is fixed to 0.1 for the all 
models. Three sets of model profiles, each with a different
mass-accretion rate, are presented here: 
(1) $\dot{M}_{\mathrm{acc}}=10^{-7}\,M_{\sun}\,\mathrm{yr^{-1}}$ (Figure~\ref{fig:atlas01}), 
(2) $\dot{M}_{\mathrm{acc}}=10^{-8}\,M_{\sun}\,\mathrm{yr^{-1}}$ (Figure~\ref{fig:atlas01}), and
(3) $\dot{M}_{\mathrm{acc}}=10^{-7}\,M_{\sun}\,\mathrm{yr^{-1}}$
(Figure~\ref{fig:atlas01}).   The temperature of the magnetosphere
used for (1), (2) and (3) are $7500\Kelvin$, $8500\Kelvin$ and
$9500\Kelvin$ respectively. The profiles are presented for the
inclination angles of $10^\circ$, $55^\circ$, and $80^\circ$. 
The disc-wind geometry used to compute these profiles are same as in
section~\ref{sub:result-discwind-hybrid}.

@@

  \caption{An H$\alpha$ atlas of the disc-wind-magnetosphere hybrid model with
  $\dot{M_{\mathrm{acc}}} =10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-6}\,M_{\sun}\,\mathrm{yr{-1}}$. 
  Profiles are presented
  for inclinations of $10^\circ$ (top figure), $55^\circ$ (middle figure), and $80^\circ$ (bottom figure).
  For each inclination, 12 profiles are given, for three different values of the wind temperature, and four
  different values of the wind acceleration parameter $\beta$.}

@@

  \caption{As for Figure~\ref{fig:atlas01} except
  $\dot{M_{\mathrm{acc}}} =10^{-8}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-7}\,M_{\sun}\,\mathrm{yr{-1}}$. }

@@

  \caption{As for Figure~\ref{fig:atlas01} except
    $\dot{M_{\mathrm{acc}}} =10^{-9}\,M_{\sun}\,\mathrm{yr{-1}}$ and
  $\dot{M_{\mathrm{wind}}}=10^{-10}\,M_{\sun}\,\mathrm{yr{-1}}$. }
